---
title: "R Notebook"
output:
  pdf_document: default
  html_document:
    df_print: paged
---


```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
```



### Introduction

Unmanned Aerial Vehicles (UAVs) play an important role in agricultural research because they facilitate high-throughput phenotyping (HTP).The ability to identify cotton plant height and boll count across a field can serve as an important tool in predicting plant growth and yield. In order to capture a three-dimensional (3D) view of field plots, which is believed to be helpful in estimating yield and crop development parameters, sensors mounted on UAVs must have access to a view of the ground. However, cotton planted in solid rows can obscure this view. Canopy closure prevents sensors from measuring plant architecture and boll-loads three dimensionally from the midgrowing
season until the crop is defoliated. Therefore, this project was initiated to compare solid vs. skip-row planting patterns in terms of predicting yield and fiber quality since skip rows would allow UAV sensors to capture more accurate 3D data from plots. The purposes of this project were to 
(1) compare the accuracy of UAV-derived data from different row patterns
(2) evaluate the ability of UAVs to predict plant yield and 
(3) characterize genotype x row pattern interaction and how location and year affect that interaction.


#### Objective 1: Accuracy between different row patterns

The height measured by UAV and human was compared.
```{r}
df <- read.csv("/Users/wenzhuowu/Desktop/tamu-project/height1.csv", header=TRUE)
head(df)
```

```{r}
lm1 = lm(UAV_h~Manual_h, data = df[df$row_pattern=='solid',]) #Create the linear regression
summary(lm1) 
library(ggplot2)


ggplot(df[df$row_pattern=='solid',],aes(Manual_h, UAV_h)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE, color='turquoise4') +
  theme_minimal() +
  labs(x='Manually meaasured plant height', y='UAV derived plant height', title='Solid row pattern') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold')) 
```
```{r}
lm2 = lm(UAV_h~Manual_h, data = df[df$row_pattern=='skip',]) #Create the linear regression
summary(lm2) 
ggplot(df[df$row_pattern=='skip',],aes(Manual_h, UAV_h)) +
  geom_point() +
  geom_smooth(method='lm', se=FALSE, color='turquoise4') +
  theme_minimal() +
  labs(x='Manually meaasured plant height', y='UAV derived plant height', title='Skip row pattern') +
  theme(plot.title = element_text(hjust=0.5, size=20, face='bold')) 
```

The skip-row planting pattern provided a more accurate plant height (R2 0.97) with lower levels of error (RMSE 3.557) compared to data
collected from the solid-row pattern (R2 0.77; RMSE 8.942 ). Row pattern may have an impact on the accuracy of plant height model based on UAV images.

#### Objective 2: Yield prediction

Plant height (ph), canopy colume (cv), canopy cover (cc), vegetation index NDVI, ExG were generated from UAV images across multiple dates. Boll count and boll area were processed based on the images taken the day before harvest.

```{r, results='hide'}
library(plyr)
library(readr)
library(dplyr)
library(glmnet)
library(ggplot2)
library(tidyverse)
library(caret)
library(leaps)
library(MASS)

```

```{r}
dat <- read.csv("/Users/wenzhuowu/Desktop/tamu-project/timeline.csv", header=TRUE)
dat = na.omit(dat)
head(dat)
```



Data partition

```{r}
set.seed(100) 

index = sample(1:nrow(dat), 0.8*nrow(dat)) 

train = dat[index,] # Create the training data 
test = dat[-index,] # Create the test data
```

Scaling the Numeric Features

```{r}
cols = colnames(dat)[-1]
pre_proc_val <- preProcess(train[,cols], method = c("center", "scale"))

train[,cols] = predict(pre_proc_val, train[,cols])
test[,cols] = predict(pre_proc_val, test[,cols])

```

stepwise regression


```{r}
# Set seed for reproducibility
set.seed(123)
# Set up repeated k-fold cross-validation
train.control <- trainControl(method = "cv", number = 10)
# Train the model
step.model <- train(Yield.per.row ~., data = dat,
                    method = "leapBackward", 
                    tuneGrid = data.frame(nvmax = 1:34),
                    trControl = train.control
                    )
step.model$results
```
```{r}
step.model$bestTune
```
```{r,results="hide"}
summary(step.model$finalModel)
```
```{r}
m1 = lm(Yield.per.row ~ CV0730+ NDVI0730, 
   data = dat)
summary (m1)
```

For stepwise regrssion, the RSE is 1.349 and R2 is 53.89 percent.



ridge 

```{r}
cols_reg = colnames(dat)
dummies <- dummyVars(Yield.per.row ~ ., data = dat[,cols_reg])
train_dummies = predict(dummies, newdata = train[,cols_reg])
test_dummies = predict(dummies, newdata = test[,cols_reg])
```

```{r}
x = as.matrix(train_dummies)
y_train = train$Yield.per.row

x_test = as.matrix(test_dummies)
y_test = test$Yield.per.row

lambdas <- 10^seq(2, -3, by = -.1)
ridge_reg = glmnet(x, y_train, nlambda = 25, alpha = 0, family = 'gaussian', lambda = lambdas)
summary(ridge_reg)
```


```{r}
cv_ridge <- cv.glmnet(x, y_train, alpha = 0, lambda = lambdas)
optimal_lambda <- cv_ridge$lambda.min
optimal_lambda
```

The optimal lambda value comes out to be 0.01 and will be used to build the ridge regression model. 
We  also create a function for calculating and printing the results, which is done with the eval_results() function in the code below. The next step is to use the predict function to generate predictions on the train and test data. Finally, we use the eval_results function to calculate and print the evaluation metrics.

```{r}
# Compute R^2 from true and predicted values
eval_results <- function(true, predicted, df) {
  SSE <- sum((predicted - true)^2)
  SST <- sum((true - mean(true))^2)
  R_square <- 1 - SSE / SST
  RMSE = sqrt(SSE/nrow(df))

  
  # Model performance metrics
data.frame(
  RMSE = RMSE,
  Rsquare = R_square
)
  
}

# Prediction and evaluation on train data
predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
eval_results(y_train, predictions_train, train)

# Prediction and evaluation on test data
predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
eval_results(y_test, predictions_test, test)
```

The above output shows that the RMSE and R-squared values for the ridge regression model on the training data are 0.0240 and 99.98 percent, respectively. For the test data, the results for these metrics are 0.6105 and 91.48 percent, respectively. 


Lasso

```{r}
lambdas <- 10^seq(2, -3, by = -.1)

# Setting alpha = 1 implements lasso regression
lasso_reg <- cv.glmnet(x, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 5)

# Best 
lambda_best <- lasso_reg$lambda.min 
lambda_best
```

The optimal lambda value is 0.001, we train the lasso model in the first line of code below. The second through fifth lines of code generate the predictions and print the evaluation metrics for both the training and test datasets.

```{r}
lasso_model <- glmnet(x, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)

predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
eval_results(y_train, predictions_train, train)

predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
eval_results(y_test, predictions_test, test)
```
The above output shows that the RMSE and R-squared values on the training data are 0.0025 and 99.99 percent, respectively. The results on the test data are 0.0064 and 99.99 percent, respectively. Lasso regression can also be used for feature selection because the coeﬃcients of less important features are reduced to zero.

```{r, results="hide"}
# Set training control
train_cont <- trainControl(method = "repeatedcv",
                              number = 10,
                              repeats = 5,
                              search = "random",
                              verboseIter = TRUE)

# Train the model
elastic_reg <- train(Yield.per.row ~ .,
                           data = train,
                           method = "glmnet",
                           preProcess = c("center", "scale"),
                           tuneLength = 10,
                           trControl = train_cont)

```

```{r}
# Best tuning parameter
elastic_reg$bestTune
```

After we have trained the model, the optimal alpha is 0.86 and lambda is 0.0021.


```{r}
# Make predictions on training set
predictions_train <- predict(elastic_reg, x)
eval_results(y_train, predictions_train, train) 

# Make predictions on test set
predictions_test <- predict(elastic_reg, x_test)
eval_results(y_test, predictions_test, test)
```

The above output shows that the RMSE and R-squared values for the elastic net regression model on the training data are 0.0563 and 99.99 percent, respectively. The results for these metrics on the test data are 0.0694 and 99.89 percent, respectively.





